---
title: Latest learnings
author: Francis Barton
date: '2020-04-06'
categories:
  - stringr
  - twirl
slug: latest-learnings
output:
  html_document:
    keep_md: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, progress = FALSE, verbose = FALSE)

library(dplyr)
library(janitor)
library(jsonlite)
library(purrr)
library(stringr)

```

### This week in R I learned that ...

Passing two vectors to `str_replace_all` as the `pattern` and `replacement` arguments does not work.
Not even if you use `purrr::map2`.

You have to pass them as a named list, where the pattern is the name and the replacement is the vector.

Let's have a look...

```{r}

# import LA - LAD - NUTS3 lookup table
la_nuts_lookup_json <- "https://opendata.arcgis.com/datasets/9b4c94e915c844adb11e15a4b1e1294d_0.geojson"

# get sample data

la_nuts_lookup <- fromJSON(la_nuts_lookup_json) %>% 
  pluck("features", "properties") %>% 
  clean_names()

df_sample <- la_nuts_lookup %>% 
  select(1:2, 5:6) %>% 
  dplyr::filter(str_detect(nuts318nm, "^(Bristol|Bath|Herefordshire|Kingston|Kensington|Glo|York)"))

head(df_sample)


nuts_patterns <- df_sample %>% 
  pull(nuts318nm) %>% 
  unique() %>% 
  str_subset(., "^(Bristol|Kingston|Bath|Hereford|Kensington)")

nuts_repl <- c("Herefordshire", "Bristol", "BANES/N Somerset/S Glos", "Kensington & Chelsea and H'smith & Fulham", "Kingston upon Hull")

```

```{r attempt1}

df_sample %>% 
  str_replace_all(nuts_patterns, nuts_repl)


```

This error message:

> In `stri_replace_all_regex(string, pattern, fix_replacement(replacement)`,  :
  longer object length is not a multiple of shorter object length
  
really confused me, because I was certain that my two vectors nuts_patterns and nuts_repl were the same length.

```{r attempt2}

df_sample %>% 
  mutate_at(vars(nuts318nm), ~ str_replace_all(., nuts_patterns, nuts_repl))

```

It didn't even work with `map2`!:

```{r attempt3}

# df_sample %>% 
  # mutate_at(vars(nuts318nm), ~ map2(nuts_patterns, nuts_repl, ~ str_replace_all(., .x, .y)))

```

Then a bit of searching led me to a post which mentioned [this in the stringr reference](https://stringr.tidyverse.org/reference/str_replace.html):


```
# If you want to apply multiple patterns and replacements to the same
# string, pass a named vector to pattern.
fruits %>%
  str_c(collapse = "---") %>%
  str_replace_all(c("one" = "1", "two" = "2", "three" = "3"))
#> [1] "1 apple---2 pears---3 bananas"
```

Which meant I just had to do this:

```{r}
names(nuts_repl) <- nuts_patterns

df_sample %>% 
  mutate_at(vars(nuts318nm), ~ str_replace_all(., nuts_repl))
```





