---
title: 'My latest problem: building a list with accumulate'
author: Francis Barton
date: '2020-07-05'
slug: building-with-accumulate
tags:
  - purrr
  - learning
---

### Another R learning curve

I just realised that yet another job I thought would be simple for me to do in R, turns out to be quite the learning curve.
<!--more-->

Here's the challenge - and I get that it seems simple, and I feel dim for not immediately knowing how to do it:

* starting with a "seed" `x` (could be a number, string, data frame), apply a function f(x) to return `y`.
* Now add `y` to a vector or list such that you get `list(x, y)`
* Now run `f()` on `y` to return `z` and add that to the list - so the list is growing incrementally through the repeated application of `f(x)`
* Stop growing the list at some limit (list length, value of `x`, whatever) and complete, returning the whole list.

I got really stuck doing this on a more complicated project, so I thought, "let's work this out using a really simple example until I understand what to do."

I have a feeling that `purrr::accumulate` is something that might be useful?
But I have not yet succeeded in making that work for anything.
Perhaps it is a sledgehammer to crack a nut.
Let's see if I can get my head around what's needed - the basic structure of the task.

### Attempt 1

So I started off with a little function:

```{r}
square_it <- function(x) {
  y <- x^2
  
  out <- list(x, y)

  # trying to do a recursive thing
  if (y < 1e5) {
    square_it(y)
  }
  else {
    return(out)
  }
}

squares <- square_it(2)
squares
```

Hmm, OK, so that sort of worked.
It did the recursive thing and stopped (roughly) when I expected it to.
It returns the last two numbers x and y, not the whole list.
This is where I think I need `purrr::accumulate`?

But first I have a few tweaks to try.

### Attempt 2

```{r}
square_it <- function(x) {
  y <- x^2
  c(x, y)

  # recursion!
  while (y < 1e4) {
    square_it(y)
  }
}

# don't run!
# square_it(2)
```

Hmmm, that *didn't* work.
It went off on one and needed me to manually break it.

While I think of it, it might be worth me just reminding myself of something similar that I *do* know how to do:
a linear progression along a vector using `purrr::map`

```{r}
library(dplyr, warn.conflicts = FALSE)
library(purrr)

raise_it <- function(n, x) {
  `^`(x, n)
}

# 1:6 %>% 
#   map_int( ~ raise_it(., 2)) # oops, weird

1:6 %>% 
  map_dbl( ~ raise_it(., 2))

```

If I know the vector to map along, then it's fine.
Where I'm stuck is where the vector/list gets built by the function as we go.
It feels a bit like this:

<div style="width:100%;height:0;padding-bottom:80%;position:relative;"><iframe src="https://giphy.com/embed/3oz8xxYqZDkPLfH7yM" width="100%" height="100%" style="position:absolute" frameBorder="0" class="giphy-embed" allowFullScreen></iframe></div><p><a href="https://giphy.com/gifs/aardman-wallace-and-gromit-train-chase-feathers-mcgraw-3oz8xxYqZDkPLfH7yM">The classic Gromit and Feathers on the train scene</a></p>

One of the confusions/uncertainties I am trying to get my head around is what needs to be *inside* the function and what can sit outside, like making or manipulating the vector/list, accumulating or mapping the function, and so on.
OK, so if I found someone else with this problem, what would I advise them to do?
I'd say start small and simple.
Write a minimal function that works.
Then use it for something small.
Go from there.
Backtrack when it breaks.

### Attempt 3

```{r}
square_it <- function(x) {
  y <- x^2
  return(c(x, y))
}

square_it(2)

# ok that works

square_it <- function(x) {
  y <- tail(x)^2
  return(c(x, y))
}

square_it(2)
square_it(square_it(2))

# hmmm weird was not expecting length == 4
square_it(c(2, 4)) # check

# aha/d'oh!

square_it <- function(x) {
  y <- tail(x, 1)^2
  return(c(x, y))
}

square_it(2)
square_it(square_it(2))

square_it <- function(x) {
  y <- tail(x, 1)^2
  z <- c(x, y)
  return(z)
  
  while (y < 1e6) {
    square_it(z)
  }
}

# nope
# square_it(2)


square_it <- function(x) {
  y <- tail(x, 1)^2
  z <- c(x, y)
  
  while (y < 1e6) {
    square_it(z)
  }
  
  return(z)
}

# still doesn't work
# square_it(2)


square_it <- function(x) {
  y <- tail(x, 1)^2
  c(x, y)
}

l <- square_it(2)
l <- c(l, square_it(l))
l # not quite

# try again
l <- square_it(2)
l <- square_it(l)
l

# see, this works but I can't work out how to formularise it!!!
# it must be something very easy!
l <- square_it(2)
l <- square_it(l)
l <- square_it(l)
l <- square_it(l)
l


l <- square_it(2)
while (tail(l, 1) < 1e7) {
  l <- square_it(l)
}
l


```

At last it worked!

Now can I rewrite it to use a list rather than a vector (useful for my actual project)?


```{r}

square_it <- function(x) {
  y <- dplyr::last(x)^2
  list(x, y)
}


l <- 2
while (dplyr::last(l) < 1e7) {
  l <- square_it(l)
}
l
```

Oooh. Mmmm. Interesting.

```{r}
# need to `flatten` the list each time
square_it <- function(x) {
  y <- dplyr::last(x)^2
  list(x, y) %>% 
    purrr::flatten()
}

# now this works beautfully
l <- 2
while (dplyr::last(l) < 1e7) {
  l <- square_it(l)
}
l

# and now as a control function???
build_squares <- function(x) {
  while (dplyr::last(x) < 1e7) {
    x <- square_it(x)
  }
  return(x)
}

squares <- build_squares(2)
squares
```































